---
title: "Lyrics data quality, matching, and repetition analysis by decade"
author: "Olivia Hong"
date: "4/30/2026"
output: html_document
---

## Setup

```{r install-packages, eval=FALSE}
install.packages(c("rvest", "httr", "dplyr", "stringr", "tidyr", "tibble", "ggplot2", "forcats", "tidytext"))
```

```{r setup, message=FALSE, warning=FALSE}
library(rvest)
library(httr)
library(dplyr)
library(stringr)
library(tidyr)
library(tibble)
library(ggplot2)
library(forcats)
library(tidytext)
```

## 1. Song URLs and matching table

The original file stores URLs in three separate vectors: 2010s, 2000s, and 90s. This section combines them into one table and creates a readable song label from each Genius URL. The decade column is the key matching variable used throughout the analysis.

```{r song-urls}
song_urls_2010s <- c(
  "https://genius.com/Kendrick-lamar-alright-lyrics",
"https://genius.com/Grimes-oblivion-lyrics",
"https://genius.com/Robyn-dancing-on-my-own-lyrics",
"https://genius.com/Beyonce-formation-lyrics",
"https://genius.com/Frank-ocean-thinkin-bout-you-lyrics",
"https://genius.com/Azealia-banks-212-lyrics",
"https://genius.com/Mitski-your-best-american-girl-lyrics",
"https://genius.com/Solange-cranes-in-the-sky-lyrics",
"https://genius.com/Lana-del-rey-video-games-lyrics",
"https://genius.com/Lil-uzi-vert-xo-tour-llif3-lyrics",
"https://genius.com/Bill-callahan-riding-for-the-feeling-lyrics",
"https://genius.com/Sky-ferreira-everything-is-embarrassing-lyrics",
"https://genius.com/Chief-keef-i-dont-like-lyrics",
"https://genius.com/Angel-olsen-shut-up-kiss-me-lyrics",
"https://genius.com/Rihanna-work-lyrics",
"https://genius.com/The-1975-love-it-if-we-made-it-lyrics",
"https://genius.com/Jamie-xx-gosh-lyrics",
"https://genius.com/Kanye-west-runaway-lyrics",
"https://genius.com/Vampire-weekend-hannah-hunt-lyrics",
"https://genius.com/M83-midnight-city-lyrics",
"https://genius.com/Drake-hold-on-were-going-home-lyrics",
"https://genius.com/Anohni-drone-bomb-me-lyrics",
"https://genius.com/Rosalia-malamente-capr-augurio-lyrics",
"https://genius.com/Todd-terje-inspector-norse-lyrics",
"https://genius.com/Fka-twigs-cellophane-lyrics",
"https://genius.com/Bon-iver-holocene-lyrics",
"https://genius.com/Mia-bad-girls-lyrics",
"https://genius.com/Beyonce-countdown-lyrics",
"https://genius.com/Joanna-newsom-good-intentions-paving-co-lyrics",
"https://genius.com/Cardi-b-bodak-yellow-lyrics",
"https://genius.com/Ariel-pinks-haunted-graffiti-round-and-round-lyrics",
"https://genius.com/Fiona-apple-every-single-night-lyrics",
"https://genius.com/Lcd-soundsystem-dance-yrself-clean-lyrics",
"https://genius.com/Future-islands-seasons-waiting-on-you-lyrics",
"https://genius.com/Rae-sremmurd-black-beatles-lyrics",
"https://genius.com/Carly-rae-jepsen-run-away-with-me-lyrics",
"https://genius.com/Jai-paul-btstu-edit-lyrics",
"https://genius.com/Kanye-west-monster-lyrics",
"https://genius.com/Sophie-bipp-lyrics",
"https://genius.com/Frank-ocean-nights-lyrics",
"https://genius.com/Courtney-barnett-avant-gardener-lyrics",
"https://genius.com/Sza-the-weekend-lyrics",
"https://genius.com/Lorde-the-louvre-lyrics",
"https://genius.com/Big-thief-mary-lyrics",
"https://genius.com/Robyn-call-your-girlfriend-lyrics",
"https://genius.com/The-knife-full-of-fire-lyrics",
"https://genius.com/Tame-impala-let-it-happen-lyrics",
"https://genius.com/Ariana-grande-thank-u-next-lyrics",
"https://genius.com/Gyptian-hold-yuh-lyrics",
"https://genius.com/Arcade-fire-sprawl-ii-mountains-beyond-mountains-lyrics",
"https://genius.com/Kendrick-lamar-bitch-dont-kill-my-vibe-lyrics",
"https://genius.com/Camp-cope-the-opener-lyrics",
"https://genius.com/Solange-losing-you-lyrics",
"https://genius.com/Disclosure-latch-lyrics",
"https://genius.com/Bat-for-lashes-laura-lyrics",
"https://genius.com/Vince-staples-norf-norf-lyrics",
"https://genius.com/Taylor-swift-all-too-well-lyrics",
"https://genius.com/Miguel-adorn-lyrics",
"https://genius.com/David-bowie-blackstar-lyrics",
"https://genius.com/Future-march-madness-lyrics",
"https://genius.com/Ema-california-lyrics",
"https://genius.com/Rihanna-we-found-love-lyrics",
"https://genius.com/Bobby-shmurda-hot-nigga-lyrics",
"https://genius.com/Phoebe-bridgers-motion-sickness-lyrics",
"https://genius.com/Drake-worst-behavior-lyrics",
"https://genius.com/King-krule-out-getting-ribs-lyrics",
"https://genius.com/Kacey-musgraves-slow-burn-lyrics",
"https://genius.com/Fka-twigs-two-weeks-lyrics",
"https://genius.com/Caribou-cant-do-without-you-lyrics",
"https://genius.com/Beach-house-zebra-lyrics",
"https://genius.com/Adele-someone-like-you-lyrics",
"https://genius.com/Tyler-the-creator-yonkers-lyrics",
"https://genius.com/Perfume-genius-queen-lyrics",
"https://genius.com/Jack-u-where-are-u-now-lyrics",
"https://genius.com/Dj-koze-pick-up-lyrics",
"https://genius.com/The-national-bloodbuzz-ohio-lyrics",
"https://genius.com/Kanye-west-ultralight-beam-lyrics",
"https://genius.com/Jeremih-oui-lyrics",
"https://genius.com/Lana-del-rey-the-greatest-lyrics",
"https://genius.com/Blood-orange-youre-not-good-enough-lyrics",
"https://genius.com/Cass-mccombs-county-line-lyrics",
"https://genius.com/Dej-loaf-try-me-lyrics",
"https://genius.com/Deerhunter-desire-lines-lyrics",
"https://genius.com/Kelela-bank-head-extended-lyrics",
"https://genius.com/Charli-xcx-track-10-lyrics",
"https://genius.com/Destroyer-chinatown-lyrics"
)

song_urls_2000s <- c(
"https://genius.com/Jamelia-superstar-lyrics",
"https://genius.com/Spiller-groovejet-if-this-aint-love-lyrics",
"https://genius.com/Kylie-minogue-spinning-around-lyrics",
"https://genius.com/Eric-prydz-call-on-me-lyrics",
"https://genius.com/Armand-van-helden-my-my-my-original-club-mix-lyrics",
"https://genius.com/Melanie-c-i-turn-to-you-lyrics",
"https://genius.com/Starsailor-four-to-the-floor-thin-white-duke-mix-lyrics",
"https://genius.com/Kelis-milkshake-lyrics",
"https://genius.com/Vitamin-c-graduation-friends-forever-lyrics",
"https://genius.com/Alex-lloyd-amazing-lyrics",
"https://genius.com/Kasey-chambers-not-pretty-enough-lyrics",
"https://genius.com/Stacie-orrico-stuck-lyrics",
"https://genius.com/Amanda-perez-angel-lyrics",
"https://genius.com/Craig-david-walking-away-lyrics",
"https://genius.com/Garbage-cherry-lips-go-baby-go-lyrics",
"https://genius.com/Kt-tunstall-suddenly-i-see-lyrics",
"https://genius.com/Geri-halliwell-its-raining-men-lyrics",
"https://genius.com/Kevin-lyttle-turn-me-on-lyrics",
"https://genius.com/Eskimo-joe-black-fingernails-red-wine-lyrics",
"https://genius.com/Nickelback-how-you-remind-me-lyrics",
"https://genius.com/Thirty-seconds-to-mars-the-kill-rebirth-lyrics",
"https://genius.com/Simple-plan-welcome-to-my-life-lyrics",
"https://genius.com/Youth-group-forever-young-lyrics",
"https://genius.com/The-cat-empire-hello-lyrics",
"https://genius.com/Slinkee-minx-summer-rain-lyrics",
"https://genius.com/Bob-sinclar-love-generation-lyrics",
"https://genius.com/Moby-we-are-all-made-of-stars-lyrics",
"https://genius.com/Jewel-intuition-lyrics",
"https://genius.com/Amiel-lovesong-radio-edit-lyrics",
"https://genius.com/Michael-buble-home-lyrics",
"https://genius.com/Evermore-light-surrounding-you-lyrics",
"https://genius.com/Nerd-she-wants-to-move-native-tongues-remix-lyrics",
"https://genius.com/Gym-class-heroes-clothes-off-lyrics",
"https://genius.com/Basement-jaxx-good-luck-lyrics",
"https://genius.com/Dht-listen-to-your-heart-furious-f-ez-radio-edit-lyrics",
"https://genius.com/Scooter-the-logical-song-lyrics",
"https://genius.com/Infernal-from-paris-to-berlin-lyrics",
"https://genius.com/Flo-rida-low-lyrics",
"https://genius.com/Estelle-american-boy-lyrics",
"https://genius.com/Sean-paul-temperature-lyrics",
"https://genius.com/Katy-perry-i-kissed-a-girl-lyrics",
"https://genius.com/Lily-allen-the-fear-lyrics",
"https://genius.com/Gabriella-cilmi-sweet-about-me-lyrics",
"https://genius.com/Coldplay-viva-la-vida-lyrics",
"https://genius.com/Empire-of-the-sun-walking-on-a-dream-lyrics",
"https://genius.com/Jason-mraz-im-yours-lyrics",
"https://genius.com/Eric-hutchinson-rock-and-roll-lyrics",
"https://genius.com/The-last-goodnight-pictures-of-you-lyrics",
"https://genius.com/End-of-fashion-o-yeah-lyrics",
"https://genius.com/3oh-3-donttrustme-lyrics",
"https://genius.com/Panic-at-the-disco-i-write-sins-not-tragedies-lyrics",
"https://genius.com/Operator-please-just-a-song-about-ping-pong-lyrics",
"https://genius.com/The-veronicas-untouched-lyrics",
"https://genius.com/Thirsty-merc-20-good-reasons-lyrics",
"https://genius.com/Finger-eleven-paralyzer-lyrics")

song_urls_90s <- c(
"https://genius.com/Pavement-gold-soundz-lyrics",
"https://genius.com/Pulp-common-people-lyrics",
"https://genius.com/Dr-dre-nuthin-but-a-g-thang-lyrics",
"https://genius.com/Radiohead-paranoid-android-lyrics",
"https://genius.com/Wu-tang-clan-protect-ya-neck-lyrics",
"https://genius.com/My-bloody-valentine-only-shallow-lyrics",
"https://genius.com/Neutral-milk-hotel-holland-1945-lyrics",
"https://genius.com/Aaliyah-are-you-that-somebody-lyrics",
"https://genius.com/Beck-loser-lyrics",
"https://genius.com/Weezer-say-it-aint-so-lyrics",
"https://genius.com/Bjork-hyper-ballad-lyrics",
"https://genius.com/Aphex-twin-windowlicker-lyrics",
"https://genius.com/Nirvana-smells-like-teen-spirit-lyrics",
"https://genius.com/The-notorious-big-juicy-lyrics",
"https://genius.com/Depeche-mode-enjoy-the-silence-lyrics",
"https://genius.com/Outkast-spottieottiedopaliscious-lyrics",
"https://genius.com/Belle-and-sebastian-the-state-i-am-in-lyrics",
"https://genius.com/Daft-punk-da-funk-lyrics",
"https://genius.com/Mazzy-star-fade-into-you-lyrics",
"https://genius.com/Dj-shadow-midnight-in-a-perfect-world-lyrics",
"https://genius.com/Smashing-pumpkins-1979-lyrics",
"https://genius.com/The-breeders-cannonball-lyrics",
"https://genius.com/Spiritualized-ladies-and-gentlemen-we-are-floating-in-space-lyrics",
"https://genius.com/Built-to-spill-car-lyrics",
"https://genius.com/Mobb-deep-shook-ones-part-ii-lyrics",
"https://genius.com/Blur-girls-and-boys-lyrics",
"https://genius.com/Elliott-smith-needle-in-the-hay-lyrics",
"https://genius.com/Nas-it-aint-hard-to-tell-lyrics",
"https://genius.com/The-verve-bitter-sweet-symphony-lyrics",
"https://genius.com/The-flaming-lips-race-for-the-prize-lyrics",
"https://genius.com/Underworld-born-slippy-nuxx-lyrics",
"https://genius.com/Bonnie-prince-billy-i-see-a-darkness-lyrics",
"https://genius.com/Missy-elliott-the-rain-supa-dupa-fly-lyrics",
"https://genius.com/New-order-regret-lyrics",
"https://genius.com/Pete-rock-and-cl-smooth-they-reminisce-over-you-troy-lyrics",
"https://genius.com/Guided-by-voices-game-of-pricks-lyrics",
"https://genius.com/Sinead-oconnor-nothing-compares-2-u-lyrics",
"https://genius.com/A-tribe-called-quest-check-the-rhime-lyrics",
"https://genius.com/Beastie-boys-sabotage-lyrics",
"https://genius.com/The-orb-little-fluffy-clouds-lyrics",
"https://genius.com/The-pharcyde-passin-me-by-lyrics",
"https://genius.com/Nine-inch-nails-closer-lyrics",
"https://genius.com/The-chemical-brothers-setting-sun-lyrics",
"https://genius.com/Massive-attack-unfinished-sympathy-lyrics",
"https://genius.com/Geto-boys-mind-playing-tricks-on-me-lyrics",
"https://genius.com/Stardust-music-sounds-better-with-you-lyrics",
"https://genius.com/Pj-harvey-rid-of-me-lyrics",
"https://genius.com/Foo-fighters-everlong-lyrics",
"https://genius.com/Yo-la-tengo-autumn-sweater-lyrics",
"https://genius.com/Oasis-live-forever-lyrics",
"https://genius.com/Public-enemy-brothers-gonna-work-it-out-lyrics",
"https://genius.com/The-stone-roses-i-wanna-be-adored-lyrics",
"https://genius.com/De-la-soul-a-roller-skating-jam-named-saturdays-lyrics",
"https://genius.com/Portishead-glory-box-lyrics",
"https://genius.com/Ghostface-killah-daytona-500-lyrics",
"https://genius.com/Godspeed-you-black-emperor-moya-lyrics",
"https://genius.com/The-beta-band-dry-the-rain-lyrics",
"https://genius.com/Fugazi-repeater-lyrics",
"https://genius.com/Deee-lite-groove-is-in-the-heart-lyrics",
"https://genius.com/Sonic-youth-100-lyrics",
"https://genius.com/Pixies-alec-eiffel-lyrics",
"https://genius.com/Jay-z-dead-presidents-ii-lyrics",
"https://genius.com/Modest-mouse-trailer-trash-lyrics",
"https://genius.com/Dismemberment-plan-the-city-lyrics",
"https://genius.com/Mary-j-blige-real-love-lyrics",
"https://genius.com/The-cardigans-lovefool-lyrics",
"https://genius.com/Bikini-kill-rebel-girl-lyrics",
"https://genius.com/Saint-etienne-avenue-lyrics")

song_data <- tibble(
  url = c(song_urls_90s, song_urls_2000s, song_urls_2010s),
  decade = factor(
    c(
      rep("90s", length(song_urls_90s)),
      rep("2000s", length(song_urls_2000s)),
      rep("2010s", length(song_urls_2010s))
    ),
    levels = c("90s", "2000s", "2010s")
  )
) |>
  mutate(
    song_slug = url |>
      str_remove("^https://genius\\.com/") |>
      str_remove("-lyrics$"),
    song_label = song_slug |>
      str_replace_all("-", " ") |>
      str_squish() |>
      str_to_title()
  )

song_data
```

```{r songs-per-decade}
song_data |>
  count(decade, name = "number_of_songs")
```

## 2. Scraping lyrics

This function uses a browser-style user agent, checks for failed web responses, keeps lyric line breaks, removes section labels such as `[Chorus]`, and returns `NA` if usable lyrics are not found.

```{r scrape-lyrics}
get_lyrics <- function(url) {
  tryCatch({
    response <- httr::GET(
      url,
      httr::user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36")
    )
    
    if (httr::status_code(response) >= 400) {
      return(NA_character_)
    }
    
    page <- httr::content(response, as = "text", encoding = "UTF-8") |>
      rvest::read_html()
    
    lyrics_nodes <- rvest::html_elements(
      page,
      "[data-lyrics-container='true'], div[class*='Lyrics__Container']"
    )
    
    if (length(lyrics_nodes) == 0) {
      return(NA_character_)
    }
    
    lyrics <- lyrics_nodes |>
      rvest::html_text2() |>
      paste(collapse = "\n")
    
    lyrics <- lyrics |>
      stringr::str_remove_all("\\[.*?\\]") |>
      stringr::str_remove_all("(?i)translations.*") |>
      stringr::str_remove_all("(?i)contributors.*") |>
      stringr::str_replace_all("([a-z])([A-Z])", "\\1 \\2") |>
      stringr::str_replace_all("([0-9])([A-Za-z])", "\\1 \\2") |>
      stringr::str_replace_all("([A-Za-z])([0-9])", "\\1 \\2") |>
      stringr::str_replace_all("[ 	]+", " ") |>
      stringr::str_replace_all(" *\n+ *", "\n") |>
      stringr::str_trim()
    
    if (stringr::str_squish(lyrics) == "") {
      return(NA_character_)
    }
    
    lyrics
  }, error = function(e) {
    NA_character_
  })
}

lyrics_df <- song_data |>
  mutate(
    lyrics = sapply(url, get_lyrics, USE.NAMES = FALSE)
  ) |>
  relocate(decade, song_label, url, lyrics)

lyrics_df
```

## 3. Data quality and matching checks

These checks identify rows that may create misleading graphs: missing lyrics, very short lyrics, duplicate URLs, duplicate scraped lyrics, and missing song labels.

```{r data-quality-checks}
lyrics_quality <- lyrics_df |>
  mutate(
    lyrics_clean = replace_na(lyrics, ""),
    matched_song_info = !is.na(song_label) & stringr::str_squish(song_label) != "",
    missing_lyrics = stringr::str_squish(lyrics_clean) == "",
    character_count = nchar(lyrics_clean),
    word_count = if_else(missing_lyrics, 0L, stringr::str_count(lyrics_clean, "\\S+")),
    line_count = if_else(missing_lyrics, 0L, stringr::str_count(lyrics_clean, "\n") + 1L),
    very_short_lyrics = !missing_lyrics & word_count < 50,
    duplicate_url = duplicated(url) | duplicated(url, fromLast = TRUE),
    duplicate_lyrics = if_else(
      missing_lyrics,
      FALSE,
      duplicated(lyrics_clean) | duplicated(lyrics_clean, fromLast = TRUE)
    )
  ) |>
  select(
    decade, song_label, matched_song_info, missing_lyrics, very_short_lyrics,
    duplicate_url, duplicate_lyrics, word_count, line_count, character_count, url
  )

lyrics_quality
```

```{r problem-rows}
problem_rows <- lyrics_quality |>
  filter(
    !matched_song_info |
      missing_lyrics |
      very_short_lyrics |
      duplicate_url |
      duplicate_lyrics
  ) |>
  arrange(decade, desc(missing_lyrics), desc(very_short_lyrics), song_label)

problem_rows
```

```{r quality-summary-table}
quality_summary_by_decade <- lyrics_quality |>
  group_by(decade) |>
  summarise(
    unmatched_song_info = sum(!matched_song_info),
    missing_lyrics = sum(missing_lyrics),
    very_short_lyrics = sum(very_short_lyrics),
    duplicate_urls = sum(duplicate_url),
    duplicate_lyrics = sum(duplicate_lyrics),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = -decade,
    names_to = "quality_check",
    values_to = "count"
  )

quality_summary_by_decade
```

```{r quality-summary-graph}
if (sum(quality_summary_by_decade$count) == 0) {
  message("No quality issues were flagged, so the quality-check graph is not needed.")
} else {
  ggplot(quality_summary_by_decade, aes(x = fct_reorder(quality_check, count), y = count)) +
    geom_col() +
    coord_flip() +
    facet_wrap(~ decade) +
    labs(
      title = "Data quality checks for scraped lyric data",
      x = "Quality check",
      y = "Number of affected rows"
    )
}
```

## 4. Convert lyrics into individual words

This creates one row per word. `word_position` records the order of each word within its song.

```{r lyrics-to-words}
lyrics_to_words <- function(song_name, lyrics_text, decade_name) {
  if (is.na(lyrics_text) || stringr::str_squish(lyrics_text) == "") {
    return(tibble(decade = character(), song_label = character(), word_position = integer(), word = character()))
  }
  
  words <- lyrics_text |>
    stringr::str_to_lower() |>
    stringr::str_replace_all("[^a-z0-9\\s']", " ") |>
    stringr::str_squish() |>
    stringr::str_split("\\s+") |>
    unlist()
  
  words <- words[!is.na(words) & words != ""]
  
  tibble(
    decade = decade_name,
    song_label = song_name,
    word_position = seq_along(words),
    word = words
  )
}

word_df <- bind_rows(
  lapply(seq_len(nrow(lyrics_df)), function(i) {
    lyrics_to_words(
      song_name = lyrics_df$song_label[i],
      lyrics_text = lyrics_df$lyrics[i],
      decade_name = as.character(lyrics_df$decade[i])
    )
  })
) |>
  mutate(decade = factor(decade, levels = c("90s", "2000s", "2010s")))

word_df
```

```{r word-data-warning}
if (nrow(word_df) == 0) {
  message("No word data was created. This usually means the lyric scraping did not work. Check lyrics_quality: if missing_lyrics is TRUE, Genius did not return usable lyric text.")
}
```

## 5. Find repetitiveness by words

A more repetitive song will usually have a lower proportion of unique words. The repetition score is higher when a song repeats more vocabulary.

```{r word-repetition-score}
if (nrow(word_df) == 0) {
  word_repetition <- tibble(
    decade = character(),
    song_label = character(),
    total_words = integer(),
    unique_words = integer(),
    repeated_words = integer(),
    unique_word_ratio = numeric(),
    repetition_score = numeric()
  )
} else {
  word_repetition <- word_df |>
    group_by(decade, song_label) |>
    summarise(
      total_words = n(),
      unique_words = n_distinct(word),
      repeated_words = total_words - unique_words,
      unique_word_ratio = unique_words / total_words,
      repetition_score = 1 - unique_word_ratio,
      .groups = "drop"
    ) |>
    arrange(desc(repetition_score))
}

word_repetition
```

```{r top-word-repetition-graph, fig.height=10, fig.width=14, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(stringr)
library(tidytext)

if (nrow(word_repetition) == 0) {
  message("No repetition score graph created because there is no word data.")
} else {
  
  top_word_repetition <- word_repetition |>
    group_by(decade) |>
    slice_max(repetition_score, n = 10, with_ties = FALSE) |>
    ungroup() |>
    mutate(
      song_label_wrapped = str_wrap(song_label, width = 25)
    )
  
  ggplot(
    top_word_repetition,
    aes(
      x = reorder_within(song_label_wrapped, repetition_score, decade),
      y = repetition_score,
      fill = decade
    )
  ) +
    geom_col(show.legend = FALSE) +
    geom_text(
      aes(label = round(repetition_score, 3)),
      hjust = -0.15,
      size = 3.2
    ) +
    coord_flip(clip = "off") +
    facet_wrap(~ decade, scales = "free_y") +
    scale_x_reordered() +
    scale_fill_manual(values = c(
      "90s" = "#66C2A5",
      "2000s" = "#FC8D62",
      "2010s" = "#8DA0CB"
    )) +
    scale_y_continuous(
      expand = expansion(mult = c(0, 0.10))
    ) +
    labs(
      title = "Top word-repetitive songs",
      subtitle = "Higher values mean more repeated vocabulary",
      x = "Song",
      y = "Repetition score"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11),
      strip.text = element_text(face = "bold", size = 10),
      axis.text.y = element_text(size = 9),
      axis.text.x = element_text(size = 9),
      axis.title = element_text(size = 12)
    )
}
```


```{r decade-word-repetition}
if (nrow(word_repetition) == 0) {
  decade_word_repetition <- tibble(
    decade = character(),
    number_of_songs = integer(),
    mean_total_words = numeric(),
    mean_unique_word_ratio = numeric(),
    mean_repetition_score = numeric()
  )
} else {
  decade_word_repetition <- word_repetition |>
    group_by(decade) |>
    summarise(
      number_of_songs = n(),
      mean_total_words = mean(total_words),
      mean_unique_word_ratio = mean(unique_word_ratio),
      mean_repetition_score = mean(repetition_score),
      .groups = "drop"
    )
}

decade_word_repetition
```

```{r decade-word-repetition-graph, fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)

decade_word_repetition <- word_repetition |>
  filter(!is.na(decade)) |>
  group_by(decade) |>
  summarise(
    mean_repetition_score = mean(repetition_score, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    decade = factor(decade, levels = c("90s", "2000s", "2010s"))
  )

if (nrow(decade_word_repetition) == 0) {
  message("No decade word repetition graph created because there is no word data.")
} else {
  ggplot(decade_word_repetition, aes(x = decade, y = mean_repetition_score, fill = decade)) +
    geom_col(width = 0.7, show.legend = FALSE) +
    geom_text(
      aes(label = round(mean_repetition_score, 3)),
      vjust = -0.4,
      size = 4
    ) +
    scale_fill_manual(values = c(
      "90s" = "#66C2A5",
      "2000s" = "#FC8D62",
      "2010s" = "#8DA0CB"
    )) +
    labs(
      title = "Average word repetitiveness by decade",
      x = "Decade",
      y = "Mean repetition score"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      axis.text = element_text(size = 11),
      axis.title = element_text(size = 12)
    )
}
```

```{r most-repeated-words}
if (nrow(word_df) == 0 || nrow(word_repetition) == 0) {
  repeated_words <- tibble(decade = character(), song_label = character(), word = character(), times_used = integer())
} else {
  top_repetition_songs <- word_repetition |>
    slice_max(repetition_score, n = 10, with_ties = FALSE) |>
    select(decade, song_label)
  
  repeated_words <- word_df |>
    semi_join(top_repetition_songs, by = c("decade", "song_label")) |>
    count(decade, song_label, word, name = "times_used") |>
    filter(times_used > 1) |>
    group_by(decade, song_label) |>
    slice_max(times_used, n = 8, with_ties = FALSE) |>
    ungroup() |>
    arrange(decade, song_label, desc(times_used))
}

repeated_words
```

```{r repeated-words-graph, fig.height=10, fig.width=14}
if (nrow(repeated_words) == 0) {
  message("No repeated words graph created because repeated_words has zero rows. This usually means the lyrics were not scraped successfully.")
} else {
  
  ggplot(
    repeated_words,
    aes(
      x = fct_reorder(word, times_used),
      y = times_used,
      fill = times_used
    )
  ) +
    geom_col() +
    facet_wrap(~ song_label, scales = "free_x") +
    scale_fill_gradient(low = "skyblue", high = "darkblue") +
    labs(
      title = "Most repeated words in the 10 most repetitive songs",
      x = "Word",
      y = "Times used",
      fill = "Count"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8),
      strip.text = element_text(size = 9, face = "bold"),
      plot.title = element_text(face = "bold")
    )
}
```

## 6. Find repetitiveness by lyric lines

This checks whether entire lyric lines are repeated. This is useful because songs often repeat choruses or hooks.

```{r lyrics-to-lines}
lyrics_to_lines <- function(song_name, lyrics_text, decade_name) {
  if (is.na(lyrics_text) || stringr::str_squish(lyrics_text) == "") {
    return(tibble(decade = character(), song_label = character(), line_number = integer(), line = character()))
  }
  
  lines <- lyrics_text |>
    stringr::str_split("\n") |>
    unlist() |>
    stringr::str_to_lower() |>
    stringr::str_remove_all("\\[.*?\\]") |>
    stringr::str_replace_all("[^a-z0-9\\s']", " ") |>
    stringr::str_squish()
  
  lines <- lines[!is.na(lines) & lines != ""]
  
  tibble(
    decade = decade_name,
    song_label = song_name,
    line_number = seq_along(lines),
    line = lines
  )
}

line_df <- bind_rows(
  lapply(seq_len(nrow(lyrics_df)), function(i) {
    lyrics_to_lines(
      song_name = lyrics_df$song_label[i],
      lyrics_text = lyrics_df$lyrics[i],
      decade_name = as.character(lyrics_df$decade[i])
    )
  })
) |>
  mutate(decade = factor(decade, levels = c("90s", "2000s", "2010s")))

line_df
```

```{r line-repetition-score}
if (nrow(line_df) == 0) {
  line_repetition <- tibble(
    decade = character(),
    song_label = character(),
    total_lines = integer(),
    unique_lines = integer(),
    repeated_lines = integer(),
    line_repetition_score = numeric()
  )
} else {
  line_repetition <- line_df |>
    group_by(decade, song_label) |>
    summarise(
      total_lines = n(),
      unique_lines = n_distinct(line),
      repeated_lines = total_lines - unique_lines,
      line_repetition_score = 1 - (unique_lines / total_lines),
      .groups = "drop"
    ) |>
    arrange(desc(line_repetition_score))
}

line_repetition
```


```{r top-line-repetition-graph, fig.height=10, fig.width=12}

if (nrow(line_repetition) == 0) {
  message("No line repetition graph created because there is no lyric line data.")
} else {
  top_line_repetition <- line_repetition |>
    group_by(decade) |>
    slice_max(line_repetition_score, n = 10, with_ties = FALSE) |>
    ungroup() |>
    mutate(song_label_wrapped = str_wrap(song_label, width = 28))

  ggplot(
    top_line_repetition,
    aes(
      x = reorder_within(song_label_wrapped, line_repetition_score, decade),
      y = line_repetition_score,
      fill = decade
    )
  ) +
    geom_col() +
    coord_flip() +
    facet_wrap(~ decade, scales = "free_y") +
    scale_x_reordered() +
    scale_fill_manual(values = c(
      "90s" = "#66C2A5",
      "2000s" = "#FC8D62",
      "2010s" = "#8DA0CB"
    )) +
    labs(
      title = "Top songs with the most repeated lyric lines",
      subtitle = "Higher values mean more repeated full lines",
      x = "Song",
      y = "Line repetition score",
      fill = "Decade"
    ) +
    theme_minimal() +
    theme(
      axis.text.y = element_text(size = 8),
      axis.text.x = element_text(size = 9),
      strip.text = element_text(size = 10, face = "bold"),
      plot.title = element_text(face = "bold"),
      legend.position = "bottom"
    )
}
```

```{r decade-line-repetition}
if (nrow(line_repetition) == 0) {
  decade_line_repetition <- tibble(
    decade = character(),
    number_of_songs = integer(),
    mean_total_lines = numeric(),
    mean_line_repetition_score = numeric()
  )
} else {
  decade_line_repetition <- line_repetition |>
    group_by(decade) |>
    summarise(
      number_of_songs = n(),
      mean_total_lines = mean(total_lines),
      mean_line_repetition_score = mean(line_repetition_score),
      .groups = "drop"
    )
}

decade_line_repetition
```

```{r decade-line-repetition-graph, fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)

decade_line_repetition <- line_repetition |>
  filter(!is.na(decade)) |>
  group_by(decade) |>
  summarise(
    mean_line_repetition_score = mean(line_repetition_score, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    decade = factor(decade, levels = c("90s", "2000s", "2010s"))
  )

if (nrow(decade_line_repetition) == 0) {
  message("No decade line repetition graph created because there is no lyric line data.")
} else {
  ggplot(decade_line_repetition, aes(x = decade, y = mean_line_repetition_score, fill = decade)) +
    geom_col(width = 0.7, show.legend = FALSE) +
    geom_text(
      aes(label = round(mean_line_repetition_score, 3)),
      vjust = -0.4,
      size = 4
    ) +
    scale_fill_manual(values = c(
      "90s" = "#66C2A5",
      "2000s" = "#FC8D62",
      "2010s" = "#8DA0CB"
    )) +
    coord_cartesian(ylim = c(0, max(decade_line_repetition$mean_line_repetition_score) + 0.02)) +
    labs(
      title = "Average line repetitiveness by decade",
      x = "Decade",
      y = "Mean line repetition score"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      axis.text = element_text(size = 11),
      axis.title = element_text(size = 12)
    )
}
```

```{r most-repeated-lines}
if (nrow(line_df) == 0) {
  repeated_lines <- tibble(decade = character(), song_label = character(), line = character(), times_repeated = integer())
} else {
  repeated_lines <- line_df |>
    count(decade, song_label, line, name = "times_repeated") |>
    filter(times_repeated > 1) |>
    group_by(decade, song_label) |>
    slice_max(times_repeated, n = 5, with_ties = FALSE) |>
    ungroup() |>
    arrange(decade, song_label, desc(times_repeated))
}

repeated_lines
```

## 7. Extra graph: total words vs unique words by decade

This graph gives a broader overview of vocabulary variety across decades.

```{r total-vs-unique-by-decade}
if (nrow(word_df) == 0) {
  word_counts_decade_long <- tibble(decade = character(), measure = character(), count = integer())
} else {
  word_counts_decade_long <- word_df |>
    group_by(decade) |>
    summarise(
      total_words = n(),
      unique_words = n_distinct(word),
      .groups = "drop"
    ) |>
    pivot_longer(
      cols = c(total_words, unique_words),
      names_to = "measure",
      values_to = "count"
    )
}

word_counts_decade_long
```

```{r total-vs-unique-by-decade-graph}
if (nrow(word_counts_decade_long) == 0) {
  message("No total vs unique words graph created because there is no word data.")
} else {
  ggplot(word_counts_decade_long, aes(x = decade, y = count, fill = measure)) +
    geom_col(position = "dodge") +
    labs(
      title = "Total words compared with unique words by decade",
      x = "Decade",
      y = "Word count",
      fill = "Measure"
    )
}
```

## Short interpretation template

The data quality checks show whether the lyrics scraped successfully and whether any rows need cleaning. Rows marked as missing, very short, duplicated, or unmatched should be checked before interpreting the graphs.

The word repetition graphs compare how much vocabulary is repeated. A higher repetition score suggests that a song, or decade average, relies more heavily on repeated words. The repeated-words graph shows which specific words are driving the pattern in the most repetitive songs.

The line repetition graphs are useful for identifying repeated choruses or repeated hooks. A high line repetition score means that whole lyric lines are repeated, not just individual words.

When writing the results section, compare the decade averages first, then use the top-song graphs and repeated-word/line tables as examples that explain the pattern.
